---
title: "HR Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install all libraries 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(janitor)
library(readr)
```
                        Step 1: Collect Data

# Load both the train & test datasets
```{r}
test <- read.csv("C:/Users/prane/OneDrive/Desktop/personal/Projects/HR Analysis/test.csv", header=FALSE, stringsAsFactors=TRUE)

train <- read.csv("C:/Users/prane/OneDrive/Desktop/personal/Projects/HR Analysis/train.csv", header=FALSE, stringsAsFactors=TRUE)
```

# Understand the structure of the train data
```{r}
head(train)

colnames(test) <- unlist(test[1, ])
```
# Remove the first row from the data frame
```{r}
test <- test[-1, ]
```
# Arrange the data frame by the employee_id column
```{r}
head(test,2)
```
# Understand the structure of the test data 
```{r}
head(train)

colnames(train) <- unlist(train[1, ])
```
# Remove the first row (header row) from the data frame
```{r}
train <- train[-1, ]
head(train,2)
```
# Arrange the data frame by the employee_id column
train %>%
    arrange(employee_id)
```{r}
train %>%
    arrange(employee_id)
```

                          Step 2: Clean Data
# Remove duplicates from train data 
```{r}
train <- distinct(train)

```
                       Step 3: Analyse and Visulaize Data
```{r}
summary(train)
```
# Visualize every variable with target variable

# Education and Promotion
```{r}
train %>% 
    group_by(education, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=education, y=n, fill=is_promoted)) + geom_bar(stat='identity', position='dodge')
```

* In terms of education, promotion is focused on Bachelor's and above degree owners.

# department and promotion 
```{r}
train %>%
    group_by(department, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=department, y=n, group=is_promoted, fill=is_promoted)) + geom_bar(stat='identity', position='dodge') + theme(axis.text.x= element_text(angle=45))
  
```

# Gender and promotion 
```{r}
train %>% 
    group_by(gender, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=gender, y=n, fill=is_promoted)) + geom_bar(stat='identity', position='dodge')

train %>% 
    group_by(gender, is_promoted) %>%
    summarise(n=n()) %>%
    filter(is_promoted==1) %>%
    ggplot(aes(x=gender, y=n, fill=is_promoted)) + geom_bar(stat='identity', position='dodge')
```

* The ratio of gender among employees is nearly identical to the ratio of gender among promotions.

# recruitment_channel and promotion    
```{r}
train %>% 
    group_by(recruitment_channel, is_promoted) %>% 
    summarise(n=n()) %>% 
    ggplot(aes(x=recruitment_channel, y=n, fill=is_promoted)) + geom_bar(stat='identity')

train %>% 
    group_by(recruitment_channel, is_promoted) %>% 
    summarise(n=n()) %>% 
    filter(is_promoted==1) %>%
    ggplot(aes(x=recruitment_channel, y=n, fill=is_promoted)) + geom_bar(stat='identity')
```

* Recruitment channel is not decisive factor to be promoted.

# no_of_trainings and promotion
```{r}
train %>%
    group_by(no_of_trainings, is_promoted) %>%
    summarise(n=n()) %>%
    filter(is_promoted==1) %>%
    ggplot(aes(x=no_of_trainings, y=n, fill=is_promoted)) + geom_bar(stat='identity',position='dodge')    

# no_of_trainings and avg_training_score
train %>%
    summarise(no_of_trainings=no_of_trainings, avg_training_score=avg_training_score) %>%
    ggplot(aes(x=no_of_trainings, y=avg_training_score, group=no_of_trainings, fill=no_of_trainings)) + geom_boxplot()
    
```

* The higher you gets score in training session, The more chances you have in promotion opportunity.

# Age and promotion
```{r}

train %>%
    group_by(is_promoted) %>%
    summarise(age=age, n=n()) %>%
    arrange(age) %>%
    ggplot(aes(x=is_promoted, y=age, group=is_promoted, fill=is_promoted)) + geom_boxplot()

train %>%
    group_by(is_promoted) %>%
    summarise(age=age, n=n()) %>%
    arrange(age) %>%
    filter(is_promoted==1) %>%
    ggplot(aes(x=is_promoted, y=age, group=is_promoted, fill=is_promoted)) + geom_violin()
    

```

* Promotion is focused on early 30s, and the chance is drastically decreased for over mid 40 years employees.

# previous year rating and promotion 
```{r}
   
train %>% 
    group_by(previous_year_rating, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=is_promoted, y=n, group=previous_year_rating, fill=previous_year_rating)) + geom_bar(stat='identity', position='dodge')
    

```

* Employees who got rated 1, 2 are merely promoted, but, in terms of ratio, employees who got rated 5 show highest promotion rate.

# Length of service and promotion
```{r}

train %>%
    group_by(is_promoted) %>%
    summarise(length=length_of_service) %>%
    #filter(is_promoted==1) %>%
    ggplot(aes(x=is_promoted, y=length, group=is_promoted, fill=is_promoted)) + geom_violin()
    

    
```

* Most promotion is focused on employees under 10 year length of services.

# KPIs met >80% and promotion
```{r}
colnames(train)[11] <- 'KPImet80'

train %>%
    group_by(KPImet80, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=is_promoted, y=n, group=KPImet80, fill=KPImet80)) + geom_bar(stat='identity', position='dodge') 
    
```

* When you met KPI requirements over 80 %, you make you chance of promotion twice.

# awards_won and promotion
```{r}

colnames(train)[12] <- 'awards_won'

train %>% 
    group_by(awards_won, is_promoted) %>%
    summarise(n=n()) %>%
    ggplot(aes(x=is_promoted, y=n, group=awards_won, fill=awards_won)) + geom_bar(stat='identity', position='dodge')

```

* Awards record is not a decisive factor for employees to be promoted. 

# avg_training_score and promotion
```{r}
train %>%
    group_by(is_promoted) %>%
    summarise(avg_training_score=avg_training_score) %>%
    ggplot(aes(x=is_promoted, y=avg_training_score, group=is_promoted, fill=is_promoted)) + geom_boxplot()
    
```

* To be promoted, achieving at least over 70 training scores is recommended.

# Step 4 : Summary & Recommendation
1) Among 8.5% of promotion chances, for employees want to be promoted, as an         analyst, I recommend 

* To get Bachelor's degree or above.
* To record over 70 % of training score on your the very first training session.
* To notice sales & marketing, operations, procurement, technology, and analytics department have top 5     promotion spots.
* To remind that most promotion chances are focused on 30 years old employees, and mid-40 is age limit.
* To achieve at least 3 from previous year rating, and over 4 ratings has more chances.
* To get over 80% of rating of KPI requirements.

2) For employers want to promote competent employees in efficient system

* Award is a irrelevant factor for promotion, so I recommend that setting a new and efficient system for    awards and promotion.
* Salary is not on data. it caused a severe problem to analyze competence and compensations including       promotion and salary.
* In recruitment channel, 'Other' has too many cases. I recommend to classify it in detail.
* The less employees take training session, the more employees got promoted. I think training session is    changed into testing session. I recommend to training session must be focused on train employees, not     assess them.